---
title: "bitcoin"
output: html_document
---
### Import Libaries
```{r}
# All coin history data was sourced from [Crypto Download Data](https://www.cryptodatadownload.com/)
library(ggplot2)
library(dplyr)
install.packages("devtools")
devtools::install_github("ropensci/plotly")
```

# Read Bitcoin Historical Data
```{r}
# Read CSV data into R
btc_data <- read.csv("Coinbase_BTCUSD_d.csv", header = TRUE)
# Display the first 6 elements to ensure that the data is read
head(btc_data)
```

## Clean Data
```{r}
# Order rows by date
btc = btc_data[order(btc_data$Date),]
head(btc)
```

```{r}
# Remove Timestamp & Symbol
btc = subset(btc, select = -c(Timestamp, Symbol))
# Check
head(btc)
```
```{r}
# Convert Date factor into date format
btc$Date <- as.Date(btc$Date, format = "%Y-%m-%d")
head(btc)
```
## Display Summary
```{r}
summary(btc)
```
No irregularities to correct. The substantial variance between prices is due to Bitcoin's rapid ascent in price over the duration of the dataset. As shown in the next section.
## Evaluating Bitcoin's Historical Price
```{r}
btc %>%
  select(Date, Close, Volume.BTC) %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))
```
```{r}
# Map the correlation between Bitcoin's closing price and volume traded (in USD)
plot(btc$Close, btc$Volume.USD)
```
```{r}
# Map the correlation between Bitcoin's closing price and volume traded (in BTC)
plot(btc$Close, btc$Volume.BTC)
```
## Price By Year
### 2014
```{r}
# 2014 Price Chart & Summary
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2014-01-01") %>%
  filter(Date <= "2014-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Date >= "2014-01-01") %>%
  filter(Date <= "2014-12-31") %>%
  summary(btc)
```
### 2015
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2015-01-01") %>%
  filter(Date <= "2015-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Date >= "2015-01-01") %>%
  filter(Date <= "2015-12-31") %>%  
  summary(btc)
```
### 2016
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2016-01-01") %>%
  filter(Date <= "2016-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Date >= "2016-01-01") %>%
  filter(Date <= "2016-12-31") %>%
  summary(btc)
```
### 2017
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2017-01-01") %>%
  filter(Date <= "2017-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Date >= "2017-01-01") %>%
  filter(Date <= "2017-12-31") %>%
  summary(btc)
```
### 2018
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2018-01-01") %>%
  filter(Date <= "2018-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Date >= "2018-01-01") %>%
  filter(Date <= "2018-12-31") %>%
  summary(btc)
```
### 2019
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2019-01-01") %>%
  filter(Date <= "2019-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Date >= "2019-01-01") %>%
  filter(Date <= "2019-12-31") %>%
  summary(btc)
```
### 2020
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2020-01-01") %>%
  filter(Date <= "2020-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Date >= "2020-01-01") %>%
  filter(Date <= "2020-12-31") %>%
  summary(btc)
```
# Compare Cryptocurrencies
### Ethereum Data
```{r}
# Read CSV data into R
eth_data <- read.csv("Coinbase_ETHUSD_d.csv", header = TRUE)
# Order rows by date
eth = eth_data[order(eth_data$Date),]
# Remove Timestamp & Symbol
eth = subset(eth, select = -c(Unix.Timestamp, Symbol))
# Convert Date factor in date format
eth$Date <- as.Date(eth$Date, format = "%Y-%m-%d")
# Display the first 6 elements to ensure that the data is read properly
head(eth)
```
### Litecoin Data
```{r}
# Read CSV data into R
ltc_data <- read.csv("Coinbase_LTCUSD_d.csv", header = TRUE)
# Order rows by date
ltc = ltc_data[order(ltc_data$Date),]
# Remove Timestamp & Symbol
ltc = subset(ltc, select = -c(Unix.Timestamp, Symbol))
# Convert Date factor into date format
ltc$Date <- as.Date(ltc$Date, format = "%Y-%m-%d")
# Display the first 6 elements to ensure that the data is read properly
head(ltc)
```
### Ripple Data
```{r}
# Read CSV data into R
xrp_data <- read.csv("Bitstamp_XRPUSD_d.csv", header = TRUE)
# Order rows by date
xrp = xrp_data[order(xrp_data$Date),]
# Remove Timestamp & Symbol
xrp = subset(xrp, select = -c(Unix.Timestamp, Symbol))
# Convert Date factor into date format
xrp$Date <- as.Date(xrp$Date, format = "%Y-%m-%d")
# Display the first 6 elements to ensure that the data is read properly
head(xrp)
```
### Monero Data
```{r}
# Read CSV data into R
xmr_data <- read.csv("Poloniex_XMRUSD_d.csv", header = TRUE)
# Order rows by date
xmr = xmr_data[order(xmr_data$Date),]
# Remove Timestamp & Symbol
xmr = subset(xmr, select = -c(Unix.Timestamp, Symbol))
# Convert Date factor into date format
xmr$Date <- as.Date(xmr$Date, format = "%Y-%m-%d")
# Display the first 6 elements to ensure that the data is read properly
head(xmr)
```
### Dash Data
```{r}
# Read CSV data into R
dash_data <- read.csv("Poloniex_DASHUSD_d.csv", header = TRUE)
# Order rows by date
dash = dash_data[order(dash_data$Date),]
# Remove Timestamp & Symbol
dash = subset(dash, select = -c(Unix.Timestamp, Symbol))
# Convert Date factor into date format
dash$Date <- as.Date(dash$Date, format = "%Y-%m-%d")
# Display the first 6 elements to ensure that the data is read properly
head(dash)
```
# Compare With Traditional Investment Vehichles (Equities & bonds)
(representde by the Dow jones and US treasury bonds)









```{r}
ds <- data.frame(Date = btc$Date, btc$Price)

fig <- plot_ly(ds, x = ~Date)
fig <- fig %>% add_lines(y = ~btc$Close, name = "Apple")
fig <- fig %>% layout(
    title = "Stock Prices",
    xaxis = list(
      rangeselector = list(
        buttons = list(
          list(
            count = 3,
            label = "3 mo",
            step = "month",
            stepmode = "backward"),
          list(
            count = 6,
            label = "6 mo",
            step = "month",
            stepmode = "backward"),
          list(
            count = 1,
            label = "1 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 1,
            label = "YTD",
            step = "year",
            stepmode = "todate"),
          list(step = "all"))),

      rangeslider = list(type = "date")),

    yaxis = list(title = "Price"))

fig
```

*Evaluate Bitcoin Price Compared to the Top 5 Currencies*
```{r}
#plot_top_5_currencies()
```
