---
title: "bitcoin"
output: html_document
---
### Import Libaries
```{r}
library(ggplot2)
library(dplyr)
#library(tidyverse)
```

# Read Bitcoin Historical Data
```{r}
# Read CSV data into R
btc_data <- read.csv("Coinbase_BTCUSD_d.csv", header = TRUE)
# Display the first 6 elements to ensure that the data is read
head(btc_data)
```

## Clean Data
```{r}
# Order rows by date
btc = btc_data[order(btc_data$Date),]
head(btc)
```

```{r}
# Remove Timestamp & Symbol
btc = subset(btc, select = -c(Timestamp, Symbol))
# Check
head(btc)
```
```{r}
# Convert Date <fctr> in <date> format
btc$Date <- as.Date(btc$Date, format = "%Y-%m-%d")
head(btc)
```
```{r}
# Create Year Category
btc$Year <- format(as.Date(btc$Date, "%Y-%m-%d"), "%Y")
head(btc)
```
```{r}
# Create Month Category
btc$Month <- format(as.Date(btc$Date, "%Y-%m-%d"), "%m")
head(btc)
```
## Display Summary
```{r}
summary(btc)
```
No irregularities to correct. The substantial variance between prices is due to Bitcoin's rapid ascent in price over the duration of the dataset. As shown in the next section.
# Evaluating Bitcoin's Historical Price
```{r}
btc %>%
  select(Date, Close, Volume.BTC) %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))
```
## Price By Year
### 2014
```{r}
# 2014 Price Chart & Summary
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2014-01-01") %>%
  filter(Date <= "2014-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Year == 2014) %>%
  summary(btc)
```
### 2015
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2015-01-01") %>%
  filter(Date <= "2015-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Year == 2015) %>%
  summary(btc)
```
### 2016
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2016-01-01") %>%
  filter(Date <= "2016-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Year == 2016) %>%
  summary(btc)
```
### 2017
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2017-01-01") %>%
  filter(Date <= "2017-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Year == 2017) %>%
  summary(btc)
```
### 2018
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2018-01-01") %>%
  filter(Date <= "2018-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Year == 2018) %>%
  summary(btc)
```
### 2019
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2019-01-01") %>%
  filter(Date <= "2019-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Year == 2019) %>%
  summary(btc)
```
### 2020
```{r}
btc %>%
  select(Date, Close, Volume.USD) %>%
  filter(Date >= "2020-01-01") %>%
  filter(Date <= "2020-12-31") %>%
  ggplot()+
  geom_line(aes(x = Date, y = Close))

btc %>%
  filter(Year == 2020) %>%
  summary(btc)
```
# Compare Cryptocurrencies



# Compare With Traditional Investment Vehichles (Equities & bonds)
(representde by the Dow jones and US treasury bonds)

```{r}

```



*Evaluate Bitcoin Price Compared to the Top 5 Currencies*
```{r}
#plot_top_5_currencies()
```
